---
title: "S15: Analysis with realistic maps and SynGO gene sets using model fit of partial least square (PLS) regression"
output:
  
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

```{css, echo=FALSE}
.watch-out {
  background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}
```

| Analysis | Atlas (Number of regions) | Rdonor | Brain data                            | Gene set                | Association          | Null model type        | Aggregation method                      |
|----------|--------------------------|--------|---------------------------------------|-------------------------|-----------------------|------------------------|-----------------------------------------|
| S15     | Desikan (34)             | 0.4    | 7 realistic maps | 49 SynGO gene sets | PLSR Fit   | Competitive / Self-contained  | NA |

*Note: For comparisons, results of the S9 analysis are also included.*


## 0. Setup
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
project_path='F:/Google Drive/post-doc/vitural_histology_revisit/revision_code'

sim_res_path=sprintf('%s/results',project_path)
result.path=sprintf('%s/reports',project_path)

atlas='desikan'
rdonor='r0.4'
brain_type='real_brain'
gene_set_type='SynGO'
cor_type='pearson'
null_type_level=c('random_gene',
                   'spin_brain')
null_type_label=c('Competitive null model',
                   'Self-contained null model')
stat_level=c('mean',
            'meanabs',
            'meansqr',
            'maxmean',
            'median',
            'ks_orig',
            'ks_weighted',
            'PLSR')
stat_label=c('Mean',
            'Meanabs',
            'Meansqr',
            'Maxmean',
            'Median',
            'KS',
            'Weighted KS',
            'PLSR')
```

## 1. Load functions

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
source(sprintf('%s/functions/analysis_functions.R',project_path))
source(sprintf('%s/functions/data_functions.R',project_path))
source(sprintf('%s/functions/cor_functions.R',project_path))
```

## 2. Load Results
```{r }

# get the list of csv files
res.files=list(
  spin_brain=sprintf( '%s/Res_%s_%s_%s_%s_spin_brain_%s_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type,cor_type),
  random_gene=sprintf('%s/Res_%s_%s_%s_%s_random_gene_%s_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type,cor_type),
  plsr_spin_brain=sprintf('%s/Res_%s_%s_%s_%s_spin_brain_PLSR_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type),
  plsr_random_gene=sprintf('%s/Res_%s_%s_%s_%s_random_gene_PLSR_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type))
 
# read res.files
res.df.list=lapply(res.files, read.csv, stringsAsFactors = F)
res.df.list$random_gene=merge(res.df.list$random_gene, res.df.list$plsr_random_gene, by=c('geneSet','sim_brain'))
res.df.list$spin_brain=merge(res.df.list$spin_brain, res.df.list$plsr_spin_brain, by=c('geneSet','sim_brain'))
res.df.list$plsr_random_gene=NULL
res.df.list$plsr_spin_brain=NULL

nest_by='brain'
pvals.nested=lapply(res.df.list, get_pvals_nested, nest_by=nest_by, heat_plot=F)
```

## 3. Show SynGO gene sets
```{r }
#get order of the gene sets
gs=load_GeneSets(atlas='desikan',rdonor = 'r0.4', gs_type='SynGO')
SynGO.df=readxl::read_excel(sprintf('%s/data/GeneSets/syngo_ontologies.xlsx',project_path))
filtered_syngo=SynGO.df %>% 
              filter(`GO term ID` %in% names(gs)) %>% 
              mutate(size=map_dbl(`GO term ID`,~length(gs[[.]]))) %>% 
              select(`GO term ID`,`GO domain`,`GO term name`,size) %>%
              arrange(desc(size))
kable(filtered_syngo,caption = 'Table 1. SynGO gene sets',booktabs = T) %>% 
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 10)
```

## 4. Make heat plots
```{r,eval=FALSE,echo=T }
brainLabel=c('ADHD','ASD','BD','MDD','OCD','SCZ','PC1')
pheat.list=lapply(pvals.nested, function(x){x %>% ungroup() %>% 
                                            mutate(brainLabel=brainLabel,
                                                   pheat=pmap(list(pvals,brainLabel,list(filtered_syngo$`GO term ID`),list(stat_level),list(stat_label)),heat_plot_pvals_df), # use list to recycle fixed argument
                                            pheat_grob=map(pheat,as.grob))})
```
```{r,include=F}
brainLabel=c('ADHD','ASD','BD','MDD','OCD','SCZ','PC1')
pheat.list=lapply(pvals.nested, function(x){x %>% ungroup() %>% 
                                            mutate(brainLabel=brainLabel,
                                                   pheat=pmap(list(pvals,brainLabel,list(filtered_syngo$`GO term ID`),list(stat_level),list(stat_label)),heat_plot_pvals_df), # use list to recycle fixed argument
                                            pheat_grob=map(pheat,as.grob))})
```

Figure 1. Results of transcriptional associations with realistic brain maps identified using different null models. The brain maps include effect sizes of case-control comparisons in attention deficit hyperactivity disorder (ADHD), autism spectrum disorder (ASD), bipolar disorder (BD), major depressive Disorder (MDD), obsessive-compulsive disorder  (OCD), schizophrenia (SZC) and the first principal component (PC1) derived from the cortical thickness of 24,750 adult participants. Yellow squares indicate the significant associations that were determined using one of the following test statistics: Mean, Meanabs, Meansqr, Maxmean, Median, Sig Number, KS, Weighted KS, and compared against that from the different null models. The last column indicates the results of comparing the model fit of partial least square (PLSR) regression between the empirical and null models. 

```{r,results='asis',echo=TRUE,fig.height=15, fig.width=30}

for (ii in c(1:length(null_type_level))){
cat('<p style="text-align: center;">', null_type_label[ii],'</p> \n')
cat('\n')
grid.arrange(grobs=pheat.list[[null_type_level[ii]]]$pheat_grob, ncol=7)
cat('\n \n')
}
```

