---
title: "S9: Analysis with realistic brain maps and SynGO gene sets"
output:
  
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

```{css, echo=FALSE}
.watch-out {
  background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}
```

| Analysis | Atlas (Number of regions) | Rdonor | Brain data                            | Gene set                | Association          | Null model type        | Aggregation method                      |
|----------|--------------------------|--------|---------------------------------------|-------------------------|-----------------------|------------------------|-----------------------------------------|
| S9     | Desikan (34)             | 0.4    | 7 realistic maps | 49 SynGO gene sets | Pearson Correlation   | Competitive / Self-contained / Coexp-matched Competitive / Brain-specific Competitive | Mean, Meanabs, Meansqr, Maxmean, Median, Sig Number, KS, Weighted KS |


## 0. Setup
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
project_path='F:/Google Drive/post-doc/vitural_histology_revisit/revision_code'

sim_res_path=sprintf('%s/results',project_path)
result.path=sprintf('%s/reports',project_path)

atlas='desikan'
rdonor='r0.4'
brain_type='real_brain'
gene_set_type='SynGO'
cor_type='pearson'
null_type_level=c('random_gene',
                   'spin_brain',
                   'random_gene_coexp',
                   'random_gene_subset')
null_type_label=c('Competitive null model',
                   'Self-contained null model',
                   'Coexp-matched competitive null model',
                   'Brain-specific competitive null model')
stat_level=c('mean',
            'meanabs',
            'meansqr',
            'maxmean',
            'median',
            'sig_n',
            'ks_orig',
            'ks_weighted')
stat_label=c('Mean',
            'Meanabs',
            'Meansqr',
            'Maxmean',
            'Median',
            'Sig Number',
            'KS',
            'Weighted KS')
```

## 1. Load functions

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
source(sprintf('%s/functions/analysis_functions.R',project_path))
source(sprintf('%s/functions/data_functions.R',project_path))
source(sprintf('%s/functions/cor_functions.R',project_path))
```

## 2. Load Results
```{r }

# get the list of csv files
res.files=list(
  random_gene_coexp=sprintf( '%s/Res_%s_%s_%s_%s_random_gene_coexp_%s_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type,cor_type),
  random_gene_subset=sprintf( '%s/Res_%s_%s_%s_%s_random_gene_subset_%s_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type,cor_type),
  spin_brain=sprintf( '%s/Res_%s_%s_%s_%s_spin_brain_%s_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type,cor_type),
  random_gene=sprintf('%s/Res_%s_%s_%s_%s_random_gene_%s_sim7.csv',sim_res_path,atlas,rdonor,brain_type,gene_set_type,cor_type))
# read res.files
res.df.list=lapply(res.files, read.csv, stringsAsFactors = F)

nest_by='brain'
pvals.nested=lapply(res.df.list, get_pvals_nested, nest_by=nest_by, heat_plot=F)
```

## 3. Show SynGO gene sets
```{r }
#get order of the gene sets
gs=load_GeneSets(atlas='desikan',rdonor = 'r0.4', gs_type='SynGO')
SynGO.df=readxl::read_excel(sprintf('%s/data/GeneSets/syngo_ontologies.xlsx',project_path))
filtered_syngo=SynGO.df %>% 
              filter(`GO term ID` %in% names(gs)) %>% 
              mutate(size=map_dbl(`GO term ID`,~length(gs[[.]]))) %>% 
              select(`GO term ID`,`GO domain`,`GO term name`,size) %>%
              arrange(desc(size))
kable(filtered_syngo,caption = 'Table 1. SynGO gene sets',booktabs = T) %>% 
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 10)
```

## 4. Make heat plots
```{r,eval=FALSE,echo=T }
brainLabel=c('ADHD','ASD','BD','MDD','OCD','SCZ','PC1')
pheat.list=lapply(pvals.nested, function(x){x %>% ungroup() %>% 
                                            mutate(brainLabel=brainLabel,
                                                   pheat=pmap(list(pvals,brainLabel,list(filtered_syngo$`GO term ID`),list(stat_level),list(stat_label)),heat_plot_pvals_df), # use list to recycle fixed argument
                                            pheat_grob=map(pheat,as.grob))})
```
```{r,include=F}
brainLabel=c('ADHD','ASD','BD','MDD','OCD','SCZ','PC1')
pheat.list=lapply(pvals.nested, function(x){x %>% ungroup() %>% 
                                            mutate(brainLabel=brainLabel,
                                                   pheat=pmap(list(pvals,brainLabel,list(filtered_syngo$`GO term ID`),list(stat_level),list(stat_label)),heat_plot_pvals_df), # use list to recycle fixed argument
                                            pheat_grob=map(pheat,as.grob))})
```

Figure 1. Results of transcriptional associations with realistic brain maps identified using different null models. The brain maps include effect sizes of case-control comparisons in attention deficit hyperactivity disorder (ADHD), autism spectrum disorder (ASD), bipolar disorder (BD), major depressive Disorder (MDD), obsessive-compulsive disorder  (OCD), schizophrenia (SZC) and the first principal component (PC1) derived from the cortical thickness of 24,750 adult participants. Yellow squares indicate the significant associations that were determined using one of the following test statistics: Mean, Meanabs, Meansqr, Maxmean, Median, Sig Number, KS, Weighted KS, and compared against that from the different null models. 

```{r,results='asis',echo=TRUE,fig.height=15, fig.width=30}

for (ii in c(1:length(null_type_level))){
cat('<p style="text-align: center;">', null_type_label[ii],'</p> \n')
cat('\n')
grid.arrange(grobs=pheat.list[[null_type_level[ii]]]$pheat_grob, ncol=7)
cat('\n \n')
}
```

## 5. Distribution of background correlation and measures of bimodality
### Figure 2. Distributions of background correlations for realistic brain maps. Each panel displays a histogram of correlations between a particular brain map with the transcriptional profiles of background genes, with the dip test statistic and the distance between the positive and negative modes shown as in texts. The x-axis of each histogram represents the correlation coefficient, and the y-axis represents the count of correlations falling within each bin. The ADHD, ASD, SCZ and PC1 show the clear separation between positive and negative correlations. 

```{r,warning=F,fig.height=4, fig.width=15}
# show bg correlat
brain_info=get_brain_info(data_path=sprintf('%s/data',project_path),
                          brain_type = 'real_brain',
                          atlas = atlas,
                          rdonor = rdonor,
                          method = cor_type)


# to make density plot
gene_data=load_GeneExp(data_path=sprintf('%s/data/GeneExp',project_path),
                        atlas = atlas,
                        rdonor = rdonor)
brain_data=load_BrainDat(data_path=sprintf('%s/data/BrainDat',project_path),
                          atlas = atlas,
                          type=brain_type,
                          col_idx = 'all') 
geneList=corr_brain_gene(gene_data=gene_data, brain_data=brain_data,method=cor_type)

brain_maps = c("ADHD", "ASD", "BD", "MDD", "OCD", "SCZ", "PC1")
plist=list()
for (i in 1:7) {
  df2plot=data.frame(cor2plot=geneList[, i])
  # Create histogram plot
  p=ggplot(data = df2plot, aes(x = cor2plot)) +
    geom_histogram(bins = 50, fill = "grey") +
    xlim(c(-1.4,1.4))+
    ylim(c(0,520))+
    annotate("text", x = -1.2, y = 510, 
             label = paste("Dip test: ", round(brain_info[i, "modetest_stat"], 3)), 
             color = "red", hjust = 0) +
    annotate("text", x = -1.2, y = 480, 
             label = paste("Mode dist: ", round(brain_info[i, "pos_neg_dist"], 3)), 
             color = "red", hjust = 0)+
    ggtitle(brain_maps[i]) +
    xlab("Background correlations") +
    ylab("Count") +
    theme_minimal()+
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          plot.title = element_text(hjust = 0.5))
  
  # Add plot to list
  plist[[i]]=p
}
grid.arrange(grobs = plist, ncol = 7,
             left =textGrob("Count",gp=gpar(fontsize=12,font=2),rot=90),
             bottom=textGrob("Background correlations",gp=gpar(fontsize=12,font=2)))
```

